---
title: "Create nice coverage plots in R"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

# Introduction

The goal here is to make some nice coverage plots.

```{r,libs}
library("rtracklayer")
library("AnnotationHub")
library("Rsamtools")
```

Here are some functions.
Fillplot is a nice way to turn line graphs into filled plots.
Mysubset extracts bedgraph data to a plottabl format.

```{r,func}
# fillplot
fillplot <- function(x,y,...) {
plot(x,y,bty="n",...)
xx <- c(x, rev(x))
yy <- c(rep(0, length(x)), rev(y))
polygon(xx, yy,border=NA,...)
}

# subset granges to a plottable format
mysubset <- function(bdg,q){
  ol <- subsetByOverlaps(bdg,q)
  ol <- as.data.frame(ol)
  ol <- t(matrix((apply(ol,1,function(x) {
      x1<-x[c(1,2,6,1,3,6)]  
    } )),nrow=3))
  ol <- as.data.frame(ol)
  colnames(ol) <- c("chr","pos","score")
  ol$pos <- as.integer(ol$pos)
  ol$score <- as.integer(ol$score)
  ol
}
```

# ATAC-seq analysis

Here is the mapped read counts for library size normalisation

```{r,counts}
counts <- read.table("cnt.txt",row.names=1)
```

Here I'm loading bedgraph from MACS2 callpeak.
These bedgraph files were compressd with bgzip and indexed with tabix.

```{r,loaddat}
bdgc1 <- rtracklayer::import("Lacz-1-ATAC_out/NA_treat_pileup.bdg.gz", format="bedGraph")
bdgc2 <- rtracklayer::import("Lacz-2-ATAC_out/NA_treat_pileup.bdg.gz", format="bedGraph")
bdgc3 <- rtracklayer::import("Lacz-3-ATAC_out/NA_treat_pileup.bdg.gz", format="bedGraph")
bdgy1 <- rtracklayer::import("Yap-1-ATAC_out/NA_treat_pileup.bdg.gz", format="bedGraph")
bdgy2 <- rtracklayer::import("Yap-2-ATAC_out/NA_treat_pileup.bdg.gz", format="bedGraph")
bdgy3 <- rtracklayer::import("Yap-3-ATAC_out/NA_treat_pileup.bdg.gz", format="bedGraph")
```

## Regions of interest

Focusing on IDH2 to begin with.

```{r,roi}
reg1 <- read.table("regions/p01_ATAC_DN_peaks_genes_linked_to_an_enhancer_with_TEAD_motif.bed")
reg1 <- reg1[which(reg1$V2=="Idh2"),]
reg1$reg <- sapply(strsplit(reg1$V1,"_"),"[[",2)
reg1$chr <- sapply(strsplit(reg1$reg,":"),"[[",1)
reg1$start <- sapply(strsplit(reg1$reg,":"),"[[",2)
reg1$start <- as.integer(sapply(strsplit(reg1$start,"-"),"[[",1))
reg1$end <- as.integer(sapply(strsplit(reg1$reg,"-"),"[[",2))

reg1$estart <- sapply(strsplit(reg1$V3,":"),"[[",2)
reg1$eend <- sapply(strsplit(reg1$estart,"-"),"[[",2)
reg1$estart <- sapply(strsplit(reg1$estart,"-"),"[[",1)

q <- GRanges(seqnames=reg1$chr,
          ranges=IRanges(start = reg1$start -1000 , end = reg1$end +1000))

olc1 <- mysubset(bdgc1,q)
olc1$score <- olc1$score/counts[1,1]*1e6

olc2 <- mysubset(bdgc2,q)
olc2$score <- olc2$score/counts[2,1]*1e6

olc3 <- mysubset(bdgc3,q)
olc3$score <- olc3$score/counts[3,1]*1e6

oly1 <- mysubset(bdgy1,q)
oly1$score <- oly1$score/counts[4,1]*1e6

oly2 <- mysubset(bdgy2,q)
oly2$score <- oly2$score/counts[5,1]*1e6

oly3 <- mysubset(bdgy3,q)
oly3$score <- oly3$score/counts[6,1]*1e6
```

```{r,plot1,fig.width=8,fig.height=8}
XMAX=max( max(olc1$pos), max(olc2$pos), max(olc3$pos), max(oly1$pos), max(oly2$pos),max(oly3$pos) )
XMIN=min( min(olc1$pos), min(olc2$pos), min(olc3$pos), min(oly1$pos), min(oly2$pos),min(oly3$pos) )
YMAX=max( max(olc1$score), max(olc2$score), max(olc3$score), max(oly1$score), max(oly2$score),max(oly3$score) )
YMIN=min( min(olc1$score), min(olc2$score), min(olc3$score), min(oly1$score), min(oly2$score),min(oly3$score) )

par(mfrow=c(6,1))
par(mar=c(2,4,1,1))
fillplot(olc1$pos,olc1$score,type="l",xlab="chr",ylab="C1",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="blue",xaxt='n'); grid()
fillplot(olc2$pos,olc2$score,type="l",xlab="chr",ylab="C2",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="blue",xaxt='n'); grid()
fillplot(olc3$pos,olc3$score,type="l",xlab="chr",ylab="C3",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="blue",xaxt='n'); grid()
fillplot(oly1$pos,oly1$score,type="l",xlab="chr",ylab="Y1",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="red",xaxt='n'); grid()
fillplot(oly2$pos,oly2$score,type="l",xlab="chr",ylab="Y2",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="red",xaxt='n'); grid()
fillplot(oly3$pos,oly3$score,type="l",xlab="chr",ylab="Y3",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="red"); grid()
mtext(paste(reg1$V1,reg1$V2))
lines(c(reg1$start,reg1$end),c(0,0),col="gray",lwd=8)
lines(c(reg1$estart,reg1$eend),c(0,0),col="black",lwd=8)
```

## Peaks with smallest p-values

Peak no. 3 has a nice clear peak.

```{r,top}
# top regions
top <- read.table("edgeR_atac_cov2_LacZvsYap.tsv",header=TRUE)
top <- top[3,]
top$reg <- sapply(strsplit(top[,1],"_"),"[[",2)
top$chr <- sapply(strsplit(top$reg,":"),"[[",1)
top$start <- sapply(strsplit(top$reg,":"),"[[",2)
top$start <- as.integer(sapply(strsplit(top$start,"-"),"[[",1))
top$end <- as.integer(sapply(strsplit(top$reg,"-"),"[[",2))

q <- GRanges(seqnames=top$chr,
          ranges=IRanges(start = top$start -500 , end = top$end +500))

olc1 <- mysubset(bdgc1,q)
olc1$score <- olc1$score/counts[1,1]*1e6

olc2 <- mysubset(bdgc2,q)
olc2$score <- olc2$score/counts[2,1]*1e6

olc3 <- mysubset(bdgc3,q)
olc3$score <- olc3$score/counts[3,1]*1e6

oly1 <- mysubset(bdgy1,q)
oly1$score <- oly1$score/counts[4,1]*1e6

oly2 <- mysubset(bdgy2,q)
oly2$score <- oly2$score/counts[5,1]*1e6

oly3 <- mysubset(bdgy3,q)
oly3$score <- oly3$score/counts[6,1]*1e6
```

```{r,plot2,fig.width=8,fig.height=8}
XMAX=max( max(olc1$pos), max(olc2$pos), max(olc3$pos), max(oly1$pos), max(oly2$pos),max(oly3$pos) )
XMIN=min( min(olc1$pos), min(olc2$pos), min(olc3$pos), min(oly1$pos), min(oly2$pos),min(oly3$pos) )
YMAX=max( max(olc1$score), max(olc2$score), max(olc3$score), max(oly1$score), max(oly2$score),max(oly3$score) )
YMIN=min( min(olc1$score), min(olc2$score), min(olc3$score), min(oly1$score), min(oly2$score),min(oly3$score) )

par(mfrow=c(6,1))
par(mar=c(2,4,1,1))
fillplot(olc1$pos,olc1$score,type="l",xlab="chr",ylab="C1",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="blue",xaxt='n'); grid()
fillplot(olc2$pos,olc2$score,type="l",xlab="chr",ylab="C2",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="blue",xaxt='n'); grid()
fillplot(olc3$pos,olc3$score,type="l",xlab="chr",ylab="C3",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="blue",xaxt='n'); grid()
fillplot(oly1$pos,oly1$score,type="l",xlab="chr",ylab="Y1",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="red",xaxt='n'); grid()
fillplot(oly2$pos,oly2$score,type="l",xlab="chr",ylab="Y2",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="red",xaxt='n'); grid()
fillplot(oly3$pos,oly3$score,type="l",xlab="chr",ylab="Y3",ylim=c(YMIN,YMAX),xlim=c(XMIN,XMAX),bty="n",col="red"); grid()
mtext(paste(top$Name,"logFC:",round(top$logFC,2)))
lines(c(top$start,top$end),c(0,0),col="gray",lwd=8)
```

