---
title: "Pathway and TF analysis of the Yap KD data"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

## Introduction

Yap KD profile

* Reactome

* TFBS TSS

* TFBS ENH

# Packages 

```{r,packages}
suppressPackageStartupMessages({
    library("tidyverse")
    library("reshape2")
    library("DESeq2")
    library("gplots")
    library("fgsea")
    library("MASS")
    library("mitch")
})
```

# Import profiling data

Here is the Yap knock-down data we generated previously.

```{r,importdata}
x <- read.table("paired-edgeR.xls",header=TRUE,row.names=1)
head(x)
sig <- subset(x,adj.p.value<0.05)
plot(x$logCPM, x$logFC, col="gray", xlab="logCPM",ylab="log2 fold change")
points(sig$logCPM, sig$logFC, col="red")
```

# Gene sets

We will perform analysis using Reactome gene sets as well as Cistrome transcription factor targets (promoter and enhancer).

```{r,genesets}

download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip", destfile="ReactomePathways.gmt.zip")
unzip("ReactomePathways.gmt.zip")
reactome <- gmt_import("ReactomePathways.gmt")

tft_tss <- gmt_import("MouseTfPeaks.gmt")

tft_enh <- gmt_import("MouseTfPeaks_enh.gmt")

```

Lets look at the make up of the gene set library.
Reactome has more smaller sets while TFT targets has mostly 1000 genes.

```{r, genesets2}
length(reactome)
length(tft_tss)
length(tft_enh)

summary(unname(unlist(lapply(reactome,length))))
summary(unname(unlist(lapply(tft_tss,length))))
summary(unname(unlist(lapply(tft_enh,length))))

```

# Mitch

Here we will run the mitch package with reactome gene sets (human).
Reactome gives a broad brush analysis of major biochemical and signaling pathways.

```{r,reactome}
rownames(x) <- sapply(strsplit(rownames(x),"_"),"[[",1)

m2h <- read.table("mouse2human.txt.sort")
m2h[,1]=NULL

# import
m <- mitch_import(x=x,DEtype="edger",geneTable=m2h)

# calc
capture.output(
    res <- mitch_calc(x=m,genesets=reactome,priority="effect")
    , file = "/dev/null", append = FALSE,
    type = c("output", "message"), split = FALSE)

head(res$enrichment_result,20)

# report
unlink("yaprna_reactome.html")
capture.output(
    mitch_report(res, "yaprna_reactome.html")
    , file = "/dev/null", append = FALSE,
    type = c("output", "message"), split = FALSE)

```

Here I'm going to run TFBS enrichment analysis based on TSS binding.

```{r,tss}
# set up gene table
x <- read.table("paired-edgeR.xls",header=TRUE,row.names=1)
gt <-  data.frame(rownames(x))
gt$gname <-  sapply(strsplit(gt[,1],"_"),"[[",2)

# import
m <- mitch_import(x=x,DEtype="edger",geneTable=gt)

# calc
capture.output(
    res <- mitch_calc(x=m,genesets=tft_tss,priority="effect",minsetsize=50)
    , file = "/dev/null", append = FALSE,
    type = c("output", "message"), split = FALSE)

head(res$enrichment_result,20)

unlink("yaprna_tss.html")
capture.output(
    mitch_report(res, "yaprna_tss.html")
    , file = "/dev/null", append = FALSE,
    type = c("output", "message"), split = FALSE)

```

Here I'm going to run TFBS enrichment analysis based on enhancer binding.

```{r,enh}
# calc
capture.output(
    res <- mitch_calc(x=m,genesets=tft_enh,priority="effect",minsetsize=50)
    , file = "/dev/null", append = FALSE,
    type = c("output", "message"), split = FALSE)

head(res$enrichment_result,20)

unlink("yaprna_enh.html")
capture.output(
    mitch_report(res, "yaprna_enh.html")
    , file = "/dev/null", append = FALSE,
    type = c("output", "message"), split = FALSE)
```


